Predictors of frog survival following translocations - analyses

### 2022-11-01 code/translocation_survival_analysis.Rmd
Use unstandardized data in conditional effects plots

mcmc_areas(m1e, area_method = "equal height", prob = 0.90, pars = b1) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Model with standardized predictors & translocation_id as grouping variable",
          "Posterior distributions with medians & 90% uncertainty intervals")

@@

@@

-   Model does a good job of predicting survival based on group-level and individual-level characteristics.\
-   What explains the low predicted survival at 70370 (both translocations), and high versus low predicted survival at 70550 for 1st and 2nd translocations, respectively? Both sites seem to contain high quality habitat (i.e., high elevation).\
-   The predicted survival is based on assumption that only males are included. Run another set of models for females?

@@

@@

d1_elevz %>% 
  select(site_id, translocation_id, snowt1_z, elev_z, male_c, donor, surv) %>% 
  add_epred_draws(m1d_reduce, ndraws = 100, re_formula = NA) %>% 
  ggplot(aes(x = elev_z, y = surv)) +
    geom_line(aes(y = .epred, group = .draw), color = "blue", size = 0.2, alpha = 0.5) +
#    geom_point(aes(y = .epred)) +
  labs(x = "Elevation of recipient site (standardized)", y = "Probability of 1-year frog survival")

# d1 %>%  # Test of add_predicted_draws used for similar purpose - unsuccessful
#   select(site_id, translocation_id, snowt1_z, elev_z, male_c, donor, surv) %>% 
#   add_predicted_draws(m1d_reduce, ndraws = 100, re_formula = NA, newdata = d1_elevz) %>% View()
#   ggplot(aes(x = elev_z, y = surv)) +
#     geom_line(aes(y = .epred, group = .draw), color = "blue", size = 0.2, alpha = 0.5) +
#     geom_point(aes(y = .epred)) +
#   labs(x = "Elevation of recipient site (standardized)", y = "Probability of 1-year frog survival")
```

Demonstrate how to unstandardize effect of elevation. This can be applied to all other effect plots.

### Try unstandardizing two ways and confirm both give the same answer

# First: Just replace elev_z with elevation
set.seed(12) # set seed for comparison between the two
d1_elevz %>% 
  select(site_id, translocation_id, snowt1_z, elev_z, male_c, donor, surv, elevation) %>% 
  add_epred_draws(m1d_reduce, ndraws = 100, re_formula = NA) %>% 
  ggplot(aes(x = elevation, y = surv)) +
    geom_line(aes(y = .epred, group = .draw), color = "blue", size = 0.2, alpha = 0.5) +
#    geom_point(aes(y = .epred)) +
  labs(x = "Elevation of recipient site (standardized)", y = "Probability of 1-year frog survival")

# Second: Untransform elev_z based on mean and sd
set.seed(12) # Same seed for comparison

# Calculate mean and sd for elevation covariate
mean_elev = mean(d1$elevation)
sd_elev = sd(d1$elevation)

d1_elevz %>% 
  select(site_id, translocation_id, snowt1_z, elev_z, male_c, donor, surv, elevation) %>% 
  add_epred_draws(m1d_reduce, ndraws = 100, re_formula = NA) %>% 
  ggplot(aes(x = (elev_z * 2*sd_elev) + mean_elev, y = surv)) +
    geom_line(aes(y = .epred, group = .draw), color = "blue", size = 0.2, alpha = 0.5) +
#    geom_point(aes(y = .epred)) +
  labs(x = "Elevation of recipient site (standardized)", y = "Probability of 1-year frog survival")

These two approaches give the same answer and can be applied to the code below.

@@

d1_snowt1z %>% 
  select(site_id, translocation_id, snowt1_z, elev_z, male_c, donor, surv) %>% 
  add_epred_draws(m1d_reduce, ndraws = 100, re_formula = NA) %>% 
  ggplot(aes(x = snowt1_z, y = surv)) +
    geom_line(aes(y = .epred, group = .draw), color = "blue", size = 0.2, alpha = 0.5) +
#    geom_point(aes(y = .epred)) +
  labs(x = "Winter severity in year following translocation (standardized)", y = "Probability of 1-year frog survival") 
```

@@

d1_malec %>% 
  select(site_id, translocation_id, snowt1_z, elev_z, male_c, donor, surv) %>% 
  add_epred_draws(m1d_reduce, ndraws = 100, re_formula = NA) %>% 
  ggplot(aes(x = male_c, y = surv)) +
    geom_boxplot(aes(y = .epred, x = male_c)) +
#    geom_point(aes(y = .epred)) +
    labs(x = "Sex (centered)", y = "Probability of 1-year frog survival")
```

@@

pal <- c("#0d0887", "#f89540")
d1_elevz_malec %>% 
  select(site_id, translocation_id, snowt1_z, elev_z, male_c, donor, surv) %>% 
  add_epred_draws(m1d_reduce, ndraws = 100, re_formula = NA) %>% 
  ggplot(aes(x = elev_z, y = surv, color = male_c)) +
    geom_line(aes(y = .epred, group = paste(male_c, .draw)), size = 0.4, alpha = 0.5) +
#    geom_point(aes(y = .epred)) +
    labs(x = "Elevation of recipient site (standardized)", y = "Probability of 1-year frog survival") +
    scale_color_manual(values = pal)  + 
    theme_classic()
```

@@

This analysis explores the results shown in translocation_survival_exploratory.Rmd (code chunk name = "frog-survival-by-site").
This plot shows frog survival of all translocations grouped by site, and indicates that survival is highly variable across sites, but repeatable within sites.
This analysis quantifies the association between survival in the first translocation and survival in subsequent translocations.

>>>>

mcmc_areas(m1e, area_method = "equal height", prob = 0.90, pars = b1) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Model with standardized predictors & translocation_id/site_id as grouping variables",
          "Posterior distributions with medians & 90% uncertainty intervals")

@@

### Posterior analysis
```{r}
tidy(m1d_reduce, conf.int = TRUE, conf.level = 0.90)

b2 <- c("(Intercept)", "snowt1_z", "elev_z", "male_c0.548", "donor70567", "donor72996")

mcmc_areas(m1d_reduce, area_method = "equal height", prob = 0.90, pars = b2) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Reduced model with standardized predictors & translocation_id as grouping variable",
          "Posterior distributions with medians & 90% uncertainty intervals")
```

@@

-   Model does a good job of predicting survival based on group-level and individual-level characteristics.  
-   The predicted survival is based on assumption that only males are included. Run another set of models for females?

@@

### Values of important predictors for translocations of interest
```{r}
d1 %>% 
  filter(site_id == "70370" | site_id == "70550") %>% 
  distinct(translocation_id, elevation, snow_t1, donor)

-   This helps to explain the low predicted survival at 70370 (both translocations), and high versus low predicted survival at 70550 for 1st and 2nd translocations, respectively.  
  -   Without translocation_id as group-level effect, survival from two translocations at 70550 are not predicted well and both are predicted to have similar (and moderate) frog survival (based on high elevation, low snow_t1, and donor = 72996 for both translocations). Therefore, the group-level effect is responsible for good prediction of survival at this site. 
  -   Low survival at 70370 in both translocations is predicted relatively well by models with and without group-level effect. Low predicted survival can be attributed to high snow_t1 and donor = 72996 in both translocations (although high elevation of site should have positive effect).

@@

# use untransformed predictor (elevation) to make plot easier to interpret  
d1_elevz %>% 
  select(site_id, translocation_id, snowt1_z, elev_z, elevation, male_c, donor, surv) %>% 
  add_epred_draws(m1d_reduce, ndraws = 100, re_formula = NA) %>% 
  ggplot(aes(x = elevation, y = surv)) +  #   "ggplot(aes(x = (elev_z * 2*sd_elev) + mean_elev, y = surv)) +" gives identical result
    geom_line(aes(y = .epred, group = .draw), color = "blue", size = 0.2, alpha = 0.5) +
    # geom_point(aes(y = .epred)) +
  labs(x = "Elevation of recipient site (meters)", y = "Probability of 1-year frog survival")
```

@@

d1_snowt1z %>% 
  select(site_id, translocation_id, snowt1_z, snow_t1, elev_z, male_c, donor, surv) %>% 
  add_epred_draws(m1d_reduce, ndraws = 100, re_formula = NA) %>% 
  ggplot(aes(x = snow_t1, y = surv)) +
    geom_line(aes(y = .epred, group = .draw), color = "blue", size = 0.2, alpha = 0.5) +
#    geom_point(aes(y = .epred)) +
  labs(x = "Winter severity in year following translocation (% of average snow depth)", y = "Probability of 1-year frog survival") 
```

@@

d1_malec %>% 
  select(site_id, translocation_id, snowt1_z, elev_z, male_c, sex, donor, surv) %>% 
  add_epred_draws(m1d_reduce, ndraws = 100, re_formula = NA) %>% 
  ggplot(aes(x = male, y = surv)) +
    geom_boxplot(aes(y = .epred, x = sex)) +
#    geom_point(aes(y = .epred)) +
    labs(x = "Sex", y = "Probability of 1-year frog survival")
```

@@

pal <- c("#0d0887", "#f89540")
d1_elevz_malec %>% 
  select(site_id, translocation_id, snowt1_z, elev_z, elevation, male_c, sex, donor, surv) %>% 
  add_epred_draws(m1d_reduce, ndraws = 100, re_formula = NA) %>% 
  ggplot(aes(x = elevation, y = surv, color = sex)) +
    geom_line(aes(y = .epred, group = paste(sex, .draw)), size = 0.4, alpha = 0.5) +
#    geom_point(aes(y = .epred)) +
    labs(x = "Elevation of recipient site (meters)", y = "Probability of 1-year frog survival") +
    scale_color_manual(values = pal) + 
    theme_classic()
```

@@

This analysis explores the results shown in translocation_survival_exploratory.Rmd (code chunk name = "frog-survival-by-site").
This plot shows frog survival of all translocations grouped by site, and indicates that survival is highly variable across sites, but repeatable within sites. The following analysis quantifies the association between survival in the first translocation and survival in subsequent translocations.
